### 2016-10-25 paper.tex
initial commit (mostly WAC-X version)

% File clic2015.tex
% June 2015

%% Based on the style files for CLiC-IT-2014, which were, in turn,
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith
% File clic2015.tex
% June 2015

%% Based on the style files for CLiC-IT-2014, which were, in turn,
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage[a4paper]{geometry}
\usepackage{clic2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

%\setlength\titlebox{5cm}
\setlength\titlebox{6.5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\usepackage[noinline,nomargin,noindex,draft]{fixme}
\fxsetup{theme=color,mode=multiuser}
\FXRegisterAuthor{n}{note}{NOTE} 
%\nnote[inline]{this is how to make regular notes...}
%\nerror[margin]{} or \fxerror[inline]{reserved for final paper}

\newcommand\BibTeX{B{\sc ib}\TeX}
\usepackage{xspace}
\newcommand\wtv{\texttt{w2v}\xspace}

\newcommand{\mytitle}{bot.zen @ EVALITA 2016 - A minimally-deep learning PoS-tagger \\
    (trained for Italian CMC)}
\newcommand{\mypdftitle}{\mytitle}
\newcommand{\mypdfsubject}{}
\newcommand{\mystitle}{bot.zen - A minimally-deep learning PoS-tagger}
\newcommand{\mykeywords}{PoS tagger, word empbeddings, word2vec, neural
    network, RNN, LSTM, EVALITA 2016, PoSTWITA}
\newcommand{\mypdfkeywords}{\mykeywords}

%%%%% hyper & options
%\definecolor{darkred}{rgb}{0.5,0,0}
%\definecolor{darkgreen}{rgb}{0,0.5,0}
%\definecolor{darkblue}{rgb}{0,0,0.5}
%
%\usepackage[final=true, pdfstartview=FitH]{hyperref}
%\hypersetup{
%    pdftitle={\mypdftitle},
%    pdfauthor={egon w. stemle <egon.stemle@eurac.edu>},
%    pdfsubject={\mypdfsubject},
%    pdfkeywords={\mypdfkeywords},
%    setpagesize=true,
%    colorlinks=true,
%    linkcolor=darkred,
%    urlcolor=darkblue,
%    citecolor=darkgreen
%}

\title{\mytitle}

\author{Egon W.~Stemle \\
    Institute for Specialised Communication and Multilingualism \\
    EURAC Research \\
    Bolzano/Bozen, Italy
    {\tt egon.stemle@eurac.edu}}

\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle

\begin{abstract} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{English.}  
This article describes the system that participated in the \emph{POS tagging
for Italian Social Media Texts} (PoSTWITA) task of the 5th periodic evaluation
campaign of Natural Language Processing (NLP) and speech tools for the Italian
language EVALITA 2016.

The system continues the work of \newcite{stemle:2016:WAC-X} with minor
modifications and different data sets.  
It combines a small assertion of trending techniques, which implement matured
methods, from NLP and ML to achieve competitive results on PoS tagging of
Italian Twitter texts; in particular, the system uses word embeddings and
character-level representations of word beginnings and endings in a LSTM RNN
architecture.  
Labelled data (Italian UD corpus, DiDi and PoSTWITA) and unlabbelled data
(Italian C4Corpus and PAIS{\`A}) were used for training.

The system is available under the APLv2 open-source license.
\end{abstract}

\begin{abstract-alt}
    \textrm{\bf{Italiano.}} Foo Bar
\end{abstract-alt}

%\footnote{\url{http://corpora.ficlit.unibo.it/PoSTWITA/}}
%\footnote{\url{http://www.evalita.it/2016}}

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:intro}

Part-of-speech (PoS) tagging is an essential processing stage for virtually all
NLP applications.
Subsequent tasks, like parsing, named-entity recognition, event
detection, and machine translation, often utilise PoS tags, and benefit
(directly or indirectly) from accurate tag sequences.

Actual work on PoS tagging, meanwhile, mainly concentrated on standardized
texts for many years, and frequent phenomena in computer-mediated communication
(CMC) and Web corpora such as emoticons, acronyms, interaction words, iteration
of letters, graphostylistics, shortenings, addressing terms, spelling
variations, and boilerplate%
~\cite{androutsopoulos2007,BernardiniBaroniEvert2008,beisswenger2013} still
deteriorate the performance of PoS-taggers%
~\cite{giesbrecht2009,baldwin-EtAl:2013:IJCNLP}.

On the other hand, the interest in automatic evaluation of social media texts,
in particular for microblogging texts such as tweets, has been growing
considerably, and specialised tools for Twitter data have become available for
different languages. 
But Italian completely lacks such resources, both regarding annotated corpora
and specific PoS-tagging tools.\footnote{http://www.evalita.it/2016/tasks/postwita}
To this end, the \emph{POS tagging for Italian Social Media Texts} (PoSTWITA)
task was proposed for EVALITA 2016 concerning the domain adaptation of
PoS-taggers to Twitter texts.
The systems were evaluated by the organisers on unlabelled but pre-tokenised
data (4759 tokens in 301 tweets).

Our system combined \texttt{word2vec} (\wtv) word embeddings (WEs) with a
single-layer Long Short Term Memory (LSTM) recurrent neural network (RNN)
architecture.
The sequence of unlabelled \wtv representations of words is accompanied by
the sequence of n-grams of the word beginnings and endings, and is fed into the
RNN which in turn predicts PoS labels.

The paper is organised as follows: We present our system design in
Section~\ref{sec:design}, the implementation in
Section~\ref{sec:implementation}, and its evaluation in
Section~\ref{sec:casestudies}. 
Section~\ref{sec:conclusion} concludes with an outlook on possible
implementation improvements.

\section{Design} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:design}

Overall, our design takes inspiration from as far back as~\newcite{Benello1989}
who used four preceding words and one following word in a feed-forward neural
network with backpropagation for PoS tagging, builds upon the strong foundation
laid down by~\newcite{collobert:2011b} for a NN architecture and learning
algorithm that can be applied to various natural language processing tasks, and
ultimately is a variation of
% achieved an accuracy of $0.95$ on the Brown Corpus.
~\newcite{santos2014learning} who trained a NN for PoS tagging, with 
character-level and WE representations of words. 
%On the Wall Street Journal (WSJ) portion of the Penn Treebank they achieve
%accuracy scores of $0.9732$.

Also note that an earlier version of the system was used in
\newcite{stemle:2016:WAC-X} to participate in the \emph{EmpiriST 2015 shared
task on automatic linguistic annotation of computer-mediated communication /
social media}~\cite{empirist:2016:WAC-X}.

\subsection{Word Embeddings} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Recently, state-of-the-art results on various linguistic tasks were
accomplished by architectures using neural-network based WEs.
\newcite{baroni-dinu-kruszewski:2014:P14-1} conducted a set of
experiments comparing the popular
\wtv~\cite{DBLP:journals/corr/abs-1301-3781,arXiv:1310.4546}
implementation for creating WEs to other distributional methods with
state-of-the-art results across various (semantic) tasks. 
These results suggest that the word embeddings substantially
outperform the other architectures on semantic similarity and analogy
detection tasks.
Subsequently,~\newcite{TACL570} conducted a comprehensive set of
experiments and comparisons that suggest that much of the improved results are
due to the system design and parameter optimizations, rather than the selected
method.  
They conclude that "there does not seem to be a consistent significant
advantage to one approach over the other".

Word embeddings provide high-quality low dimensional vector representations of
words from large corpora of unlabelled data, and the representations, typically
computed using NNs, encode many linguistic regularities and
patterns~\cite{arXiv:1310.4546}.

\subsection{Character-Level Sub-Word Information} %%%%%%%%%%%%%%%%%%%%%%%%%%%%

The morphology of a word is opaque to WEs, and the relatedness of
the meaning of a lemma's different word forms, i.e.~its different string
representations, is \emph{not} systematically encoded. 
This means that in morphologically rich languages with long-tailed frequency
distributions, even some WE representations for word forms of common
lemmata may become very poor~\cite{DBLP:journals/corr/KimJSR15}.

We agree with~\newcite{santos2014learning}
and~\newcite{DBLP:journals/corr/KimJSR15} that sub-word information is very
important for PoS tagging, and therefore we augment the WE representations with
character-level representations of the word beginnings and endings; thereby, we
also stay language agnostic---at least, as much as possible---by avoiding the
need for, often language specific, morphological pre-processing.

\subsection{Recurrent Neural Network Layer} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Language Models are a central part of NLP.
They are used to place distributions over word sequences that encode systematic
structural properties of the sample of linguistic content they are built from,
and can then be used on novel content, e.g.~to rank it or predict some feature
on it. 
For a detailed overview on language modelling research see~\newcite{mikolov2012}.

A straight-forward approach to incorporate WEs into feature-based
language models is to use the embeddings' vector representations as
features.\footnote{For an overview see,
e.g.~\newcite{Turian:2010:WRS:1858681.1858721}.}
Having said that, WEs are also used in neural network architectures, where they
constitute (part of) the input to the network.

Neural networks (NNs) consist of a large number of simple, highly interconnected
processing nodes in an architecture loosely inspired by the structure of the
cerebral cortex of the brain~\cite{oreilly2000}.
The nodes receive weighted inputs through these connections and \emph{fire}
according to their individual thresholds of their shared activation function.
A firing node passes on an activation to all successive connected nodes.
During learning the input is propagated through the network and the output is
compared to the desired output. 
Then, the weights of the connections (and the thresholds) are adjusted
step-wise so as to more closely resemble a configuration that would produce the
desired output.
After all input cases have been presented, the process typically starts over
again, and the output values will usually be closer to the correct values.

RNNs are NNs where the connections between the elements are directed cycles,
i.e.~the networks have loops, and this enables them to model sequential
dependencies of the input.
However, regular RNNs have fundamental difficulties learning long-term
dependencies, and special kinds of RNNs need to be used~\cite{Hochreiter1991}; 
a very popular kind is the so called long short-term memory (LSTM)
network proposed by~\newcite{Hochreiter:1997:LSM:1246443.1246450}.

Overall, with this design we not only benefit from available labelled data but
also from available general and domain-specific unlabelled data.

\section{Implementation} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:implementation}

We maintain the implementation in a source code repository at
\url{https://github.com/bot-zen/}.  
The version tagged as {\tt 1.1} comprises the version that was used to generate
the results submitted to the ST.

Our system feeds WEs and character-level sub-word information into a
single-layer RNN with a LSTM architecture.

\subsection{Word Embeddings} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%When computing WEs we keep in mind that \newcite{TACL570} observed that one
%specific configuration of~\wtv, namely the skip-gram model with negative
%sampling (SGNS) 
%"is a robust baseline.  While it might not be the best method for every task,
%it does not significantly underperform in any scenario. Moreover, SGNS is the
%fastest method to train, and cheapest (by far) in terms of disk space and
%memory consumption".
%Coincidentally,~\newcite{arXiv:1310.4546} also suggest to use SGNS
We incorporates \wtv's original C implementation for learning WEs%
\footnote{\url{https://code.google.com/archive/p/word2vec/}}
in an independent pre-processing step, i.e.~we pre-compute the WEs.
Then, we use gensim%
\footnote{\url{https://radimrehurek.com/gensim/}}%
, a Python tool for unsupervised semantic modelling from plain text, to load
the data, and to extract the vector representations of the embedded words as
input to our NN.

\subsection{Character-Level Sub-Word Information} %%%%%%%%%%%%%%%%%%%%%%%%%%%%

Our implementation uses a \emph{one-hot encoding} with a few additional
features for representing sub-word information.
The one-hot encoding transforms a categorical feature into a vector where
the categories are represented by equally many dimensions with binary values.
We convert a letter to lower-case and use the sets of ASCII characters, digits,
and punctuation marks as categories for the encoding.
Then, we add dimensions to represent more binary features like
\emph{'uppercase'} (was upper-case prior to conversion), \emph{'digit'} (is
digit), \emph{'punctuation'} (is punctuation mark), \emph{whitespace} (is white
space, except the new line character; note that this category is usually empty,
because we expect our tokens to \emph{not} include white space characters), and
\emph{unknown} (other characters, e.g.~diacritics).
This results in vectors with more than a single \emph{one-hot} dimension. 
%
% - alpha-characters are mapped to lower-case one-hot + 'is-uppercase'
% - digits are mapped to one-hot + 'is-digit'
% - punctuation marks are mapped to one-hot + 'is-punctuation'
% - whitespace (ecxept '\n') characters are mapped to one-hot +
% 'is-whitespace'
% - unknowns have their own one-hot
% * '\n' is treated as new-token-character

\subsection{Recurrent Neural Network Layer} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Our implementation uses Keras, a minimalist, highly modular NNs library,
written in Python and capable of running on top of either TensorFlow or
Theano~\cite{chollet2015}. 
In our case it runs on top of Theano, a Python library that allows to define,
optimize, and evaluate mathematical expressions involving multi-dimensional
arrays efficiently~\cite{Theano2016}.

The input to our network are sequences of the same length as the sentences we
process.
During training we group sentences of the same length into batches.
Each single word in the sequence is represented by its sub-word information and
two WEs that come from two sources (see Section~\ref{sec:casestudies}).
%Note that the WEs of the domain-specific corpus may have only been derived
%from a small corpus.
Unknown words, i.e.~words without a WE, are mapped to a randomly generated
vector representation once, and this representation is reused later.
In Total, each word is represented by $1,800$ features: two times $500$ (WEs),
and ten times $80$ for two 5-grams (word beginning and ending).
(If words are shorter than 5 characters their 5-grams are zero-padded.)

This sequential input is fed into a LSTM layer that, in turn, projects to a
fully connected output layer with softmax activation function.
We use categorical cross-entropy as loss function and backpropagation in
conjunction with the RMSprop optimization for learning.  
At the time of writing, this was the Keras default---or the explicitly
documented option to be used---for our type of architecture.

\section{Case Study} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:casestudies}

% For the proposed task we plan to re-use the tweets being part of the EVALITA2014 SENTIPLOC corpus: both the development and test set will be manually annotated for a global amount of 4041+1749 tweets and will be distributed as the new development set. Then a new manually annotated test set composed of 600/700 tweets will be produced using texts from the same period of time. All the annotations will be carried out by three different annotators. 
% 
% We will distribute a tokenised version of the texts in order to avoid tokenisation problems among participants and the boring problem of disappeared tweets.

We used our implementation to participate in the EmpiriST 2015 shared task.
First, we describe the corpora used for training, and then the specific system
configuration(s) for the ST.
%% and subsequently, to run experiments with different data and parameters for
%%training (see Subsection~\ref{ssec:case_postempirist}).
%Following, a short description of the data we used for training.

\subsection{Training Data for \wtv and PoS Tagging} %%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Tiger v2.2 (PoS)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{Tiger v2.2}%
\footnote{\url{http://www.ims.uni-stuttgart.de/forschung/ressourcen/korpora/tiger.html}}
is version 2.2 of the TIGERCorpus~\cite{Brants2004} containing German newspaper
texts. 
The corpus was semi-automatically PoS tagged, and is one of the standard
corpora used for German PoS tagging. 
It contains 888,238 tokens in 50,472 sentences.
For research and evaluation purposes, the TIGERCorpus can be downloaded for
free.

\subsubsection{German Wikipedia (\wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{de.wiki'15}%
\footnote{\url{http://www1.ids-mannheim.de/kl/projekte/korpora/verfuegbarkeit.html\#Download}}
are user talk pages (messages from users to users, often questions and advice),
article talk pages (questions, concerns or comments related to improving a
Wikipedia article), and article pages of the German wikipedia from 2015, made
available by the \emph{Institut f\"{u}r Deutsche
Sprache}\footnote{\url{http://www.ids-mannheim.de}}.
%% user talk pages
%% 378,576,711 toks
%% 15,343,904 s
%%
%% article talk pages
%% 446,551,142 toks 
%% 17,275,385 s
%%
%% article pages
%% 1,126,462,560 toks
%% 46,740,611 s
%%
% tokens: 378,576,711+446,551,142+1,126,462,560 = 1,951,590,413
% sentences: 15,343,904+17,275,385+46,740,611 = 79,359,900
The corpus contains 2 billion tokens 
(talk:379m, article talk:447m, article:1,1bn) 
in 79 million sentences 
(talk:15m, article talk:17m, article:47m), is well-sized for \wtv, and
also (partly) resembles or target data.
It is available under the CC BY-SA 3.0\footnote{Creative Commons
Attribution-ShareAlike 3.0 Unported, i.e.~the data can be copied and
redistributed, and adapted for any purpose, even commercially. See
\url{http://creativecommons.org/licenses/by-sa/3.0/} for more details.}
license.

\subsubsection{EmpiriST 2015 Data (PoS and \wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{subsec:empirist}
\emph{empirist}%
\footnote{\url{https://sites.google.com/site/empirist2015/home/shared-task-data}}
is the CMC and Web data made available by the organizers of the ST.
It contains data samples from different CMC genres and samples from text genres
on the Web.
The training corpus contains 10,053 tokens and was PoS tagged by two annotators
(unclear cases were decided by a third person). 
The trial corpus contains around 3,600 tokens (2,100 CMC%
\footnote{For evaluation during the development phase we used
    \emph{empirist}-trial.
    Unfortunately, we found out only later that the CMC part of the trial data
    is also part of the training data, i.e.~for the CMC data our evaluation
    data was identical with the training data.}%
, 1,500 Web) and was PoS tagged by one annotator (without systematic error
checks%
%, i.e.~the accuracy of the PoS annotations is likely lower than on the
%training data and on the ST final evaluation data
).
See~\newcite{empirist2016} for more details.

%An overview of the used training data can be found in Table~\ref{tab:corpora}.

%\begin{table}[h]
%\begin{center}
%\begin{tabular}{l|r|r}
%\hline
%\multicolumn{1}{c}{} & \multicolumn{1}{c}{Tokens} & \multicolumn{1}{c}{Sentences} \\ \hline
%empirist       & $10*10^3$     & $0.6*10^3$    \\ \hline
%Tiger v2.2     & $0.9*10^9$    & $0.05*10^6$   \\ \hline\hline
%de.wiki'15   	& $2.0*10^9$    & $79*10^6$     \\ \hline
%%C4Corpus      & $$            & $$            \\ \hline
%%DECOW14       & $11.7*10^9$   & $600*10^6$    \\ \hline
%%deTenTen	    & $19.9*10^9$   & $1300*10^6$   \\ \hline
%\end{tabular}
%\end{center}
%\caption{\label{tab:corpora}Training data used for the EmpiriST 2015 shared
%    task. The PoS tags of the upper two corpora were used for training.
%    % (upper part) and additional data for the subsequent experiments (lower
%    % part).
%} \end{table}

%\emph{C4Corpus}%
%\footnote{\url{https://github.com/dkpro/dkpro-c4corpus}} is a full documents
%German Web corpus that has been extracted from CommonCrawl, the largest
%publicly available general Web crawl to date. 
%The data is available under the CreativeCommons license
%family\footnote{\url{https://creativecommons.org/licenses/}}. 
%See Habernal~\shortcite{Habernal.et.al.2016.LREC} for details about the corpus
%construction pipeline, and other information about the corpus.
%
%\emph{DECOW14}%
%\footnote{\url{http://corporafromtheweb.org/decow14/}} is the sentence shuffled
%German Web corpus by COW created with the 2014 technology of the COW
%initiative~\cite{SchaeferBildhauer2012}.
%The data can be used for purposes of academic research.
%% $11.7*10^9$ tokens 
%% $0.6*10^9$ (624,767,747) sentences
%
%\emph{deTenTen} Web corpus.
%%DeTenTen
%% $19.9*10^9$ (19,918,263,493) tokens
%% $1.3*10^9$ (1,309,696,220) sentences
%\cite{1120431}

\subsection{EmpiriST 2015 shared task} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{ssec:case_empirist}

For the ST we used one overall configuration for the system, but we used three
different corpus configurations for training.
Consequently, we participated in the ST with three runs:
we used PoS tags from \emph{empirist} (run 1), from \emph{Tiger v2.2} (run 2),
and from both (run 3).  
For \wtv we trained a 500-dimensional skip-gram model on \emph{empirist} that
ignored all words with less than 3 occurrences within a window size of 10;
it was trained with negative sampling (value 5) and erroneously%
\footnote{%
    According to \wtv's author, technically negative sampling and
    hierarchical softmax can be combined but one should avoid combining them~%
    (see~\url{https://groups.google.com/forum/\#!topic/word2vec-toolkit/WUWad9fL0jU}).

    We had forgotten to deactivate an option in a data processing script.
}
also with hierarchical softmax.
We also trained a 500-dimensional continuous bag-of-words model on
\emph{de.wiki'15} that ignored all words with less than 25 occurrences within a
window size of 10; it was trained with negative sampling (value 3) and
erroneously also with hierarchical softmax.

The rational behind training the two models differently was that according to
\wtv author's experience%
\footnote{\url{https://groups.google.com/d/msg/word2vec-toolkit/NLvYXU99cAM/E5ld8LcDxlAJ}}
a skip-gram model "works well with small amount[s] of the training data, [and]
represents well even rare words or phrases", and a cbow model is "several times
faster to train than the skip-gram, [and has] slightly better accuracy for the
frequent words".
The other \wtv parameters were left at their default settings%
\footnote{\texttt{-sample 1e-3 -iter 5 -alpha 0.025} for skip-gram and
\texttt{-alpha 0.05} for continuous bag-of-words}.
% empirist.bin -cbow 0 -size 500 -window 10 -hs 1 -negative 5 -min-count 3 
% bigdata.bin -cbow 1 -size 500 -window 10 -hs 1 -negative 3 -min-count 25

To optimize the system's output we ran a simple grid search for three
parameters: the hidden LSTM layer's size, the dropout value for the projections
from the LSTM to the output layer during training, and the number of epochs
during training. 
The found values were size:$1024$, dropout:$0.1$, epochs:$20$.
%lstm_output_dims = [128, 256, 512, 1024*]
%dropouts = [0.1*, 0.25, 0.5, 0.75]
%nb_epochs = [5, 10, 20*]

\begin{table}[h]
\begin{center}
\begin{tabular}{l|r|r}
\hline
\multicolumn{1}{c}{} & \multicolumn{1}{c}{CMC} & \multicolumn{1}{c}{Web} \\ \hline
(1) \emph{empirist}             & $81.03$       & $86.97$ \\ \hline
(2) \emph{Tiger v2.2}           & $73.56$       & $89.73$ \\ \hline
(3) \emph{empirist}+\emph{Tiger v2.2} & $85.42$ & $90.63$ \\ \hline\hline
Winning Team                    & $87.33$       & $93.55$ \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:results}Official results of our PoS tagger for the three
runs on the EmpiriST 2015 shared task data.
} 
\end{table}

%The official results are listed in Table~\ref{tab:results}.

% 1: empirist_training (tagset:ibk)
% 2: tiger-2.2 (tagset:1999)
% 3: empirist_training + tiger-2.2 (tagset:ibk)
%Our evaluation on the EmpiriST trial data resulted in these accuracy scores:
%1:~$0.88$, 
%2:~$0.78$, 
%3:~$0.93$;
%and the overall ST evaluation was:
%1:~$0.84$,
%2:~$0.82$,
%3:~$0.88$.

%\subsection{Post-EmpiriST Experiments}
%\label{ssec:case_postempirist}
%As a post-EmpiriST addition to the results we wanted to (1)~remedy some errors
%and oversights in our initial pre-processing and training, (2)~optimize the
%learning parameters of the word embedding models and the neural network, and
%(3)~train word embedding models on other data.
%\fxnote[margin]{add (1)}
%
%The optimal learning parameters for the word embedding models were selected
%according to a comprehensive set of experiments and comparisons by
%Levy~\shortcite{TACL570}, and observations and suggestions by
%Mikolov~\shortcite{arXiv:1310.4546}. 
%For the post-EmpiriST experiments we used the skip-gram with negative-sampling
%training method (SGNS), with a negative-sampling value of $15$ (and without
%hierarchical softmax, cf.~Subsection~\ref{ssec:case_empirist}), ignored all
%words with less than $100$ occurrences within a window size of $10$, and a
%sub-sampling value of $10^{-5}$. 
%% cbow=0, negative=15, hs=0, window=10, sample=10^-5 
%% arXiv:1310.4546: -negative 15 -sample 0.00001
%% TACL570: alpha=0.75 and dyn=1-> fixed in word2vec algorithm;
%%There are quite a few differences between the skip-gram and the CBOW models.
%%However, if you have a lot of training data, their performance should be
%%comparable.%
%%\footnote{\url{https://groups.google.com/d/msg/word2vec-toolkit/NLvYXU99cAM/E5ld8LcDxlAJ}}
%
%As additional data to train word embedding models we obtained and processed
%\emph{C4Corpus}, \emph{DECOW14}, and \emph{deTenTen}.
%
%\fixme[inline]{The system is said to be trained on various widely available
%    corpora of German such as Tiger 2.2, DECOW148 etc. It is not discussed in
%    any detail and thus does not become clear why these corpora were selected
%    and especially what the systematic benefit of using these different data
%    sets was; more specifically, the paper does not systematically explore the
%    diversity of the data and their contribution to a learning improvement on
%    the system. A section with a systematic evaluation is missing and so is a
%    section with a more thorough discussion as to why this particular approach
%    was selected. It is suggested to the authors to provide more background on
%    these questions.}

\section{Conclusion \& Outlook} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:conclusion}

We presented our submission to the EmpiriST 2015 shared task, where we
participated in the PoS tagging sub-task with fair results on the CMC data and
adequate results on the Web data. 
Still, our implementation, albeit following state-of-the art designs and methods, is
quite unpolished, and can certainly gain performance with more detailed tuning.
For example, adding special sequence start and sequence stop symbols to the
input is typically done as a pre-processing step, which might improve the
results at the beginning and the end of sentences;
or we might gain some performance by adding additional hidden layers to enable
the network to learn more intermediate abstractions.
A more profound design change could also help, e.g.~Recurrent Memory Network
are a novel recurrent architecture that have been shown to outperform LSTMs on
some language modelling tasks.
%and sentence completion tasks, while also allowing insights into the data from
%a linguistic perspective~\newcite{tran-bisazza-monz:2016:N16-1}.
Finally, for learning the word embeddings we could use different corpora, or
selectively extract parts from large web-corpora resembling---as much as
possible---the type of data that is to be tagged.

%\section*{Acknowledgments} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%The acknowledgments should go immediately before the references.  Do
%not number the acknowledgments section. Do not include this section
%when submitting your paper for review.

% References %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\clearpage
%\nocite{*}
%\small
\bibliographystyle{acl}
\bibliography{paper}
%\normalsize

%\appendix %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Supplemental Material}
%\label{sec:supplemental}
%Supplement

\end{document}